---
title: "Report 01: Introduction to the package `reaov` and some functions used to draw the inference"
author: Nam-Anh Tran
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{report01NamAnh}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(reaov);
```

## Introduction

we aim to direct linear model fitting using the package `reaov`. The package involves two samples of data sets that we will use throughout the report for the illustrative purpose. Both data `mydata1` and `mydata2` are the data `gala` and `odor`, respectively, from the package `faraway`. To see the information of the data, print the code `?reaov::mydata1` or `?faraway::gala` for the first data and `?reaov::mydata2` or `?faraway::odor` for the second data.


In this report, we introduce some functions that calculate the estimates of model parameters and their inference. `LReg` is the main function with two arguments `myformula` and `mydata`, which are the formula of the model we'd like to fit and the data we use fit the linear model. This function calculates all required values, e.g. matrices $H$, $(X'X)^{-1}$, $\widehat e$, $\widehat\beta$, etc. These values will then be called directly to calculate the point and interval estimation of the model parameters without re-calculating.  

In this package, we derive the confidence intervals (CI) but p-value since the interval estimates are more appealing and can be used to make the hypothesis testing decision without calculating p-value. For the joint confidence intervals ($\ge 2$ dimentions), we calculate the CI of the corresponding statistic and make a decision. 

The object of `LReg` has 3 methods corresponding to the generic function `summary`, which follows OOP S3. The package also has 2 functions, and 1 helper function. We summarize the functions and methods in the following table: 

```{r, echo=FALSE}
dplyr::tibble(
  "method/function" = c("LReg", ".tconf", ".CIresponse", ".JCRpars", "ConTrastTest", "SubsetBetaTest", "AddClass"),
  "Calculate" = c("The main function the the package that returns all required metrics for other calculations", "Confidence intervals of model parameters", "Confidence intervals of responses and future response", "Joint confidence region for the model parameter vector", "Contrast hypothesis testing", "Calculate the sub-vector of the parameters and draw the inference for the reduced model.", " (A helper function) add a new class to the object of `LReg`"))|>
  knitr::kable()|>
  kableExtra::add_footnote("<i>Method and function are distingushed by starting with '.'</i>", notation = "symbol", escape = FALSE)
```

The three methods calculate the most appealing metrics when we fit a linear model, the confidence interval of each parameter, the confidence intervals pertaining to new observations, and the hypothesis test of the joint of all parameters. The two functions support other hypothesis tests that we may interest.  

The class `.tconf` is assigned to the `LReg` object automatically once we run `LReg`. To add other classes to the object, we use the helper function `AddClass`. Once the new class is added, the output of the generic function `summary` will be correnponding to the new class. If we want to use the previous method, we need to add that method again using `AddClass`. 

Once we run `summary`, its output will be kept track and saved (or updated if we rerun it with new defined arguments) to the object of `LReg` (this is only true for methods but functions).

## Functions and class methods mechanism

1. **The main function `LReg`:** This is the main function of the package since its output will be the first argument of most of other functions and methods. Other functions and methods extract required values from the output of `LReg` and carry out a specific task. The output involves 12 metrics: response $y$, design matrix `X`, sample size $n$, the number of covariates $p$ (excluding the intercept), the hat matrix $H$, identity matrix $I$, inverse matrix $(X'X)^{-1}$, the estimated response vector of $\widehat y$, the residual vector $\widehat e$, estimate of prameters $\widehat\beta$, estimate of variance $\sigma^2_{naive}$, and corrected estimate of variance $\sigma^2_{cor}$. For more details, run `?reaov::LReg()`. 

2. **The method `.tconf`:** This method is created automatically when we run the function `LReg`. This method has 2 arguments, where the first argument is the object of `LReg`, and the second is the level of significance $\alpha$ the default of which is $0.05$. This method function returns a tibble (data.frame) involving the point estimation, standard error, interval estimation of the parameters, and the decision of hypothesis testing for each parameter (including intercept). For more details, run `?reaov::summary.tconf`. The CI is specified by modifying the t-statistic 
$$
t_j := \frac{\widehat\beta - \beta_j}{\sigma^2_{cor}\sqrt{(X'X)^{-1}_{j,j}}},
$$
and the decision of the hypothesis is made using the CI (i.e. we reject the null hypothesis if zero falls outside of the CI).

3. **The method `.CIresponse`:** To use this method, we need to add a new method the the object of `LReg` using the helper function `AddClass`. Once the method is added, we use the generic function summary to run the method. This method has 3 arguments, the first argument, again, is the object of `LReg`. The second is new obsevation(s) that we want to estimate; the default is NULL, indicating that we want to estimate all data points that we use to fit the model. The third is the level of significance $\alpha$ whose default is $0.05$. This method returns the estimate of the mean response $X\widehat\beta$, the CI of the mean response and the CI of thr response based on the following formulae:
$$
x'_*\widehat\beta \pm t_{\alpha/2;n-p}\widehat\sigma_{cor}\sqrt{x_*'(X'X)^{-1}x_*}, \quad
x'_*\widehat\beta \pm t_{\alpha/2;n-p}\widehat\sigma_{cor}\sqrt{1+x_*'(X'X)^{-1}x_*}.
$$
Note that the second argument must in the the matrix form (not the vector even one observation is estimated). For more details, run `?summary.CIresponse`.

4. **The method `.JCRpars`:** This method is accessed only if its corresponding class is added to the object of `LReg`. The method undetakes the F test for the joint of parameters (the whole vector of parameters). The method has 2 arguments, the first is the object of `LReg`, and the second is the level of significance $\alpha$ (default is $0.05$). It returns a tibble icluding the F-statistic, the $(1-\alpha)100\%$ quantile of the F distribution, and decision of hypothesis testing, calculated using the following formula
$$
F_{statistic} =\frac{(\widehat\beta-\beta)'(X'X)(\widehat\beta -\beta)}{p\widehat\sigma_{cor}^2},
$$
Here, we test $\beta = 0$. For more details, run `?summary.JCRpars`. 

5. **The function `ConTrastTest`:** . This function undertakes the hypothesis test for the contrast vector (or matrix) R of parameters. The function has 3 arguments, the first is the object of `LReg`, The second is the matrix $R$ (the contrast matrix), and the last one is the level of significance $\alpha$ (default is 0.05). The function returns the F-statistic, $(1-\alpha)100\%$ qunatile of the F distribution, and decision of hypothesis testing, calculated using the following formula
$$
F_{statistic} = \frac{(R\widehat\beta -r)'[R(X'X)^{-1}R']^{-1}(R\widehat\beta -r)}{k\widehat\sigma^2_{cor}}.
$$
Here, we test $R\beta = 0$. For more details, run `?ContrastTest`. 

6. **The function `SubsetBetaTest`:**  This function calculates the hypothesis test for the reduced model. It has 3 arguments, the first is the object of `LReg`, the second is the vector of the ordering of parameters we'd like to omit (default is 1), and the last one is the level of significance $\alpha$ (default is 0.05). The function returns a tibble involving F-statistic, the $(1-\alpha)100%$ quantile of the F distribution and the decision of hypothesis testing. It follows the formula
$$
\frac{R(\gamma_1|\gamma_2)/k}{SS_{res}/(n-p)} \sim F_{k,n-p}, \text{ under } H_0,
$$
where $R(\gamma_1|\gamma_2)$ represents the difference between the reduced and full model, $k$ is the number of parameter we'd like to omit, and $SS_{res}$ is the residual sum of square. For more details, run `?SubsetBetaTest`. 

7. **The helper function `AddClass`** This function has two arguments, the first is the object of `LReg`, and the second is the name of the class we'd like to add to the object. 

## Illustrative examples

We start by fitting the model using the main function `LReg`; we use the data `mydata1` to fit the model. 

```{r}
fit<- LReg(Species ~ Area +Elevation+ Nearest+ Scruz+ Adjacent, mydata1)
```

when we run the function, it prints the point estimate of all parameters (including the intercept). All required values are saved in a list. We can see the name of all elements of the list as follows

```{r}
names(fit)
```

We now use to method `.tconf` to obtain the CI of each parameters

```{r}
summary(fit)
```

Note that when we run the method, its output is automatically saved to the object `fit`. We can check it as follows

```{r}
names(fit)
```

The output is saved under the name of the class, i.e. `tconf`. 

We now calculate the the CI of mean response and response using the method `CIresponse`. We first add this class to the object `fit` and run the generic function `summary`

```{r}
AddClass(fit, "CIresponse")
summary(fit)
```

We can try it again with the new observations 

```{r}
new = matrix(c(100, 300, 0.1, 2, 2, 25, 200, 2,30, 500, 100,50,1,1,600), ncol = 5)
summary(fit, x_star = new)
```

We run the method `CIresponse` twice, but only the newest is saved to the object fit. Let us extract it from the object

```{r}
fit$CIresponse
```

Next, we calculate the hypothesis test of the joint of parameters

```{r}
AddClass(fit, "JCRpars")
summary(fit)
```

Extract it from the main object

```{r}
fit$JCRpars
```

To reuse the method `tconf` we have to add it again

```{r}
AddClass(fit, "tconf")
summary(fit)
```

We Now use the function `ConTrastTest` to compare the `Area` vs. `Elevation` and `Nearest` vs. Adjacent

```{r}
R = matrix(c(0,1,-1,0,0,0,0,0,0,1,0,-1), ncol = 6, byrow = T)
ContrastTest(fit, R = R)
```

The output of the function is not saved to the main object

```{r}
names(fit)
```

Finally, We use to function `SubsetBetaTest` to test the reduced model. Let us omit the intercept and the third covariate

```{r}
SubsetBetaTest(fit)

red = c(1, 3)
SubsetBetaTest(fit, reduced = red)
```

Thus, it does not change much as we omit the intercept, but the intercept and `Elevation`. 





































