---
title: "Diagnostic Plots Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{report02}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(reaov); library(ggplot2); library(dplyr)

```

```{r, include = FALSE}

Diagnostic_Plots = function(object){
  diagnostics_tib <- as_tibble(c(list(
    e_hat = object$e_hat,
    e_hat_st = object$e_hat/diag(object$hat_mat),
    y_hat = object$y_hat),
    data.frame(object$x[,2:dim(object$x)[2]])
  ))
  x_names <- colnames(data.frame(object$x[,2:dim(object$x)[2]]))

  #make an empty list of plots, names are already filled in
  list_names <- c("residuals_v_fitted",
                  "residuals_hist",
                  paste("residuals_v_", x_names, sep = ''), #residuals_v_xi
                  "qq_standardized_residuals"
  )
  list_plots <- setNames(vector(mode = "list", length=length(list_names)), list_names)

  #residuals vs fitted values
  list_plots$residuals_v_fitted <- diagnostics_tib %>%
    ggplot(aes(x = y_hat, y = e_hat)) +
    geom_point() +
    geom_hline(yintercept = 0, linetype = "dotted", color = "red")+
    labs(title = "Residuals vs Fitted Values",
         x = "Fitted Value",
         y = "Residual")

  #residuals histogram
  list_plots$residuals_hist <- diagnostics_tib %>%
    ggplot(aes(x = e_hat)) +
    geom_histogram()+
    labs(title = "Residuals Histogram",
         x = "Residual")

  #residuals vs covariates
  for(x_name in x_names){
    list_plots[[paste("residuals_v_", x_name, sep = '')]] <- diagnostics_tib %>%
      ggplot(aes(x = .data[[x_name]], y = e_hat)) +
      geom_point() +
      geom_hline(yintercept = 0, linetype = "dotted", color = "red")+
      labs(title = paste("Residuals vs", x_name),
           x = x_name,
           y = "Residual")
  }

  #qq plot of standardized residuals vs theoretical quantile
  list_plots$qq_standardized_residuals <- diagnostics_tib %>%
    ggplot(aes(sample = e_hat_st))+
    geom_qq() +
    geom_qq_line(color = "red") +
    labs(title = "Normal QQ Plot Standardized Residuals",
         x = "Theoretical",
         y = "Standardized Residual")

  return(list_plots)
}
```

This vignette will describe how to use the Diagnostic Plots function.
We will create a model with LReg on our dataset mydata1 and proceed to go through the diagnostic plots
```{r}
modfit<- LReg(Species ~ Area +Elevation+ Nearest+ Scruz+ Adjacent, mydata1)
plots <- Diagnostic_Plots(modfit)
```
Diagnostic Plots returns a list of diagnostic plots. We will go through them.

First, we have the residuals vs fitted values plot. This plot is to visually check if there is a violation of homoskedasticity in the model. While it may be hard to see with not so many points, it looks like the homoskedasticity is holding up.
```{r}
plots$residuals_v_fitted
```


Below are residuals vs covariate plots. These check if the conditional expectation given the covariates of the residuals is indeed 0. These do not appear wildly violated, but we do not have many points and it is difficult to see with the longitudinal spread of this data.
```{r}
plots$residuals_v_Area
plots$residuals_v_Elevation
plots$residuals_v_Adjacent
plots$residuals_v_Scruz
plots$residuals_v_Nearest
```

Below is a qqplot of the standardized residuals. The theoretical quantiles being used are from the Normal distribution, so we are checking visually the assumption of normal standardized residuals. This only happens if we a linear regression function, homoskedascticity, and uncorrelated errors. As we can see below, the observed quantiles match the theoretical ones quite well.
```{r}
plots$qq_standardized_residuals
```

We can also look at the historgam of the residuals. We expect to see a normal looking distribution.
```{r}
plots$residuals_hist
```
