---
title: "Report 01: Introduction to the package `reaov`"
author: Nam-Anh Tran
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{report01NamAnh}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

we aim to direct linear model fitting using the package `reaov`. The package involves two samples of data sets that we will use throughout the report for the illustrative purpose. 

```{r setup}
library(reaov)
```

In the report 1, we introduce some functions that calculate the estimates of model parameters and their inference. `LReg` is the main function with two arguments `myformula` and `mydata`, which are the formula of the model we'd like to fit and the data that is used to fit the linear model. 

In this package, we derive the confidence intervals (CI) but p-value since the interval estimates are more appealing and can be used to make the hypothesis testing decision without calculating p-value. For the joint confidence intervals ($\ge 2$ dimentions), we calculate the CI of the corresponding statistic and make a decision. 

Function `LReg` will keep track all the required values to calculate the estimates of parameters and draw their inference. The object of `LReg` has 3 methods corresponding to the generic function `summary`, which follows OOP S3, and 1 function as follows: 

```{r, echo=FALSE}
dplyr::tibble(
  "method/function" = c(".tconf", ".CIresponse", ".JCRpars", "ConTrastTest"),
  "Calculate" = c("Confidence intervals of model parameters", "Confidence intervals of responses and future response", "Joint confidence region for the model parameter vector", "Contrast hypothesis testing"))|>
  knitr::kable()|>
  kableExtra::add_footnote("Method and function are distingushed by '.'", notation = "symbol")
```

The class `.tconf` is assigned to the `LReg` object automatically once we run `LReg`. To add other classes to the object, we use the helper function `AddClass`. Once the new class is added, the output of `summary` will pertain to the new class. If we want to use the previous method, we need to add it again using `AddClass`. 

Note than once we run `summary`, its output will be kept track and saved (or updated if we rerun it with new defined arguments) to the object of `LReg` in a list.

## Illustrative examples

We fit the linear model for the data `mydata1` (run `?reaov::mydata1` for detail) using the function `LReg` and assign it to `modelfit`. When we call `modelfit`, it prints the estimates of the intercept and coefficients corresponding to predictors.      

```{r}
modfit<- LReg(Species ~ Area +Elevation+ Nearest+ Scruz+ Adjacent, mydata1)
modfit
```

To see the list of all saved values in the object `modfit`, we can check the name of elements in the list as follows:

```{r}
names(modfit)
```

All elements in the list can be obtained using the `$` or `[[.]]` as follows

```{r}
modfit$beta_hat
modfit[c("sigma2_naive", "sigma2_cor")]
```

We obtain the CIs of coefficients using the generic function `summary`.  These CIs are calculated based on the following formula:

$$
I_{\alpha} = \widehat{\beta}_j \pm t_{\alpha/2}\widehat{\sigma}_{cor}(\mathbf{X}'\mathbf{X})^{-1/2}_{(j,j)}
$$

```{r}
summary(modfit)
```

Note that once we run the method, its output is saved to the object `modfit`. We can check it by calling the name of the list's elements

```{r}
names(modfit)
```

The output of method `.tconf` was saved as the thirteen element in the list. Also, we can check the vector of class of `modfit` as follows

```{r}
class(modfit)
```

We now calculate the CI of mean response using the method `.CIresponse`. to do this, we first add the class name to the object `modfit`

```{r}
AddClass(modfit, "CIresponse")
class(modfit)
```

We then calculate the CI using the generic function `summary`. The new  

Note that the new observations must be in matrix form (not a vector), the output involves mean response and its two CIs corresponding to mean response and future mean response. To see the arguments of this method, we can run `?summary.CIresponse`.   

```{r}
x_star = rbind(c(24,300,1,3,1.84), c(10,90,12,6,1)); x_star
summary(modfit, x_star)
```

Note that the default of argument `x_star` is predictors of all observations in the data set that is used to fit the model. 

```{r}
summary(modfit)
```

Also, the output relating to this method was saved to the object `modfit`:

```{r}
names(modfit)
modfit$CIresponse
```

Next, we calculate the joint confidence region of the vector of parameters using the method `.JCRpars`. We again add the new class to `modfit` and run the generic function `summary`. It returns the F statistic, its corresponding CI, and the decision. 

```{r}
AddClass(modfit, "JCRpars")
summary(modfit)
```

We may now be interested in the contrast hypothesis testing. This can be achieved using the function `ContrastTest`. For example, we want to compare `Area` vs. `Elevation` and `Adjacent` vs. 0.   

```{r}
contrast = rbind(c(0,1,-1,0,0,0), c(0,0,0,0,0,1))
r = c(0,0)
ContrastTest(modfit, R = contrast, r = r)
```

The output of functions is not saved as that of the methods. 






























